---
title: "BEV Adoption - Multiple Imputation and Cluster Analysis"
author: "Mark Stephens"
date: "2024-06-19"
output: html_document
---

```{r setup, include=FALSE}
# Set the root directory for all chunks
knitr::opts_knit$set(root.dir = "C:/Users/Mark Stephens/Documents/Masters Data Science work/Capstone project/Data")

knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r load packages echo=T}
pacman::p_load(dplyr, stringr,tidyverse, tidymodels,caret, future, ggplot2,mlim,janitor,stargazer,mice,cli,crayon,ucminf,nloptr,lme4,performance,devtools,reticulate,furrr,progressr,
               RCurl,jsonlite,kml3d,zoo,clusterMI,fdasrvf,timetk,xgboost,kknn,hardhat,themis,h2o,agua,reshape2)
# Download and install the H2O package for R

```
## Load data
```{r working directory}
Analysis_data <- read.csv("Analysis_df.csv")
Auxiliary_data <- read.csv("et_imputed_Auxiliary.csv")
# load R data file 

# Set dataframes
Analysis_data <- as.data.frame(Analysis_data)
Auxiliary_data <- as.data.frame(Auxiliary_data)

# Import the pickle module from Python
selected_Aux_var <- py_run_string("
import pickle
with open('selected_Aux_vars.pickle', 'rb') as file:
  data = pickle.load(file)")

#  Load pickled Auxiliary selected features
selected_Aux_var <- py$data
#print( selected_Aux_var)
# Finally, let's load H2O and start up an H2O cluster
#library(h2o)
head(Auxiliary_data)
head(Analysis_data)
```
# Further Data Cleaning

```{r Data Cleaning Analysis Data}
# Janitor 
# Clean names
Analysis_data <-
Analysis_data %>% clean_names()

Auxiliary_data <-
  Auxiliary_data %>% clean_names()

selected_Aux_var <-
  tolower(selected_Aux_var)

#rename region
Auxiliary_data <-
Auxiliary_data %>%
  rename(region = electoral_rating)

#redefine Auxiliary dataset
Auxiliary_data <-
Auxiliary_data %>% select(-c(official_name_state:log_distance_capital))


#set region as factor
Auxiliary_data <-
Auxiliary_data %>% mutate(across(c(region), factor))

# collect lga_code and lga_name
lgas <- 
  Analysis_data %>% select(lga_code, lga_name)

# drop lga_name, rooftop_solar, "households", "businesses"
Analysis_data <-
Analysis_data %>% 
  select(-c("lga_name")) %>%
  rename(state = official_name_state)

# Transform variable
Analysis_data <-
  Analysis_data %>%
  mutate(log_distance_capital = log(distance_capital))

# make greens_vote_prop as percentage
Analysis_data <- 
  Analysis_data %>% mutate(greens_vote_prop = greens_vote_prop*100)

# Remove a column
Analysis_data <-
Analysis_data %>% select(-distance_capital)

#set state as factor
Analysis_data <-
Analysis_data %>% mutate(across(c(state), factor))

#Set numeric columns to float
Analysis_data <- 
Analysis_data %>%
  mutate(across(c(higher_education,high_income:log_distance_capital),~ as.numeric(.x)))

# Define the first differencein EV_proportion
#Analysis_data <- 
#Analysis_data %>%
#  mutate(across(c(rooftop_solar,households,businesses),~ as.integer(as.numeric(.x))))

# move Target variable to end
Analysis_data <-
  Analysis_data %>%
  select(-ev_prop,ev_prop)
```

## Transform Target Variable
1. From the transformations identified in Exploratory data analysis,  the first difference of the log transformation de-trends the Target variable and gives the most linear relationship.
```{r Target variable Transformation}
# Function to calculate half the second smallest value
half_second_smallest <- function(x) {
  unique_vals <- sort(unique(x))
  if (length(unique_vals) < 2) {
    return(0) # Handle cases where there are not enough unique values
  } else {
    return(unique_vals[2] / 2)
  }
}    
ev_prop_shift <- half_second_smallest(Analysis_data$ev_prop)
#Analysis_data <- Analysis_data %>% 
#  group_by(lga_code) %>%
#  mutate(ev_prop_log = log(ev_prop + ev_prop_shift),
#         ev_prop_log_diff = c(0, diff(ev_prop_log) )) %>%
#  ungroup() %>%
#  select(-ev_prop_log)
```

## Data Segregation
Split the data into Train and Test Set based on downstream analysis forecast requirements.
```{r Data Segregation}
# Available data period
start_year <- min(unique(Analysis_data$year))
end_year <- max(unique(Analysis_data$year))

# Minimum number of years for the test data set period
test_min_years <- 5
 
# Calculate the latest year for the training data to ensure at least minimum allowable period in the test set
train_last_year <- end_year - test_min_years

# Split the data based on boolean vector
train_test_split <- Analysis_data$year <= train_last_year
training <- Analysis_data[train_test_split,]
testing <- Analysis_data[!train_test_split,]
# separate predictors and target variable
X_train <- training %>% select(-ev_prop)
y_train <- training%>% select(ev_prop)
X_test <- testing %>% select(-ev_prop)
y_test <- testing%>% select(ev_prop)

# Check the split years
print(paste("Training data includes years up to:", train_last_year))
print(paste("Dimensions of the training set is: ",sum(train_test_split)," by ",ncol(Analysis_data)))
print(paste("Dimensions of the test set is: ",sum(!train_test_split)," by ",ncol(Analysis_data)))
```

# Combine the Training data with the Auxiliary variable set
```{r merge training set and Auxiliary set}

# Combine the main training data with the auxiliary data - full variable set
data_combined <- merge(Analysis_data, Auxiliary_data, by = c("lga_code","year"), all.x = TRUE)
print(paste("dimensions of combined data set 1: ", nrow(data_combined)," by ",ncol(data_combined)))

#####Quickpred - 40 imputation model
# select variables from Auxiliary dataset from quickpred - 40

training_combined <- merge(X_train, Auxiliary_data[,c('lga_code','year',selected_Aux_var,'region')], by = c("lga_code","year"), all.x = TRUE)
print(paste("dimensions of combined training data set 2: ", nrow(training_combined)," by ",ncol(training_combined)))

testing_combined <- merge(X_test, Auxiliary_data[,c('lga_code','year',selected_Aux_var,'region')], by = c("lga_code","year"), all.x = TRUE)
print(paste("dimensions of combined training data set 2: ", nrow(testing_combined)," by ",ncol(testing_combined)))
```
### Random effect in years 
Examine intaclass correlation coefficient to see if random effects in Year accounts for observed variability. Year to year changes and unmeasured effects such as policy changes, economic conditions and natural disasters and events (pandemic)  effect variables.

```{r icc}
# missing value columns
missing_value_cols <- 
  c("higher_education",
    "population_density",
    "high_income",
    "tech_affinity",
    "greens_vote_prop")

icc_values <- list()
# Loop through each variable to fit the model and calculate ICC
for (var in missing_value_cols) {
  # Filter out missing values
  data_filtered <- Analysis_data[!is.na(Analysis_data[[var]]), ]
  # Check if the filtered data still has more than one level for the grouping factor 'year'
  if (length(unique(data_filtered$year)) > 1) {
  
    formula <- as.formula(paste(var, "~ 1 + (1 | year)"))
    model <- lmer(formula, data = data_filtered)
    
    # Check for singularity
    if (isSingular(model, tol = 1e-4)) {
      icc_values[[var]] <- NA  # Mark as NA if model is singular
      warning(paste("Model for", var, "is singular. Consider simplifying the random effects structure."))
      
    } else {
      icc_values[[var]] <- icc(model)
    }
  } else {
    icc_values[[var]] <- NA  # Not enough levels in grouping factor after filtering
  }
}

# Print ICC values
print(icc_values)
```
High Income and rooftop solar have a moderate  ICC, suggesting a non-negligible portion of their variance  is due to differences between years. Including year as a random effect can help model this variability more accurately.

### Perform MUltiple Imputation using MICE
The Auxiliary variables are combined with Analysis Variables and the d 
```{r Initialise multiple imputation for cluster analysis}
# Use a local seed
withr::local_seed(1L)
# Use {progress} package for progress bar - shows diagnostics in real time
handlers("txtprogressbar")
# Correct Column names
training_combined <-
  training_combined %>% clean_names()
# cluster id must be of type integer
training_combined$lga_code <- as.integer(training_combined$lga_code)

# Calculate the proportion of missing values in analysis feature set
total_elements <- prod(dim(Analysis_data ))  # Total number of elements in the data frame
missing_values <- sum(is.na(Analysis_data)) # Total number of missing values
proportion_missing <- missing_values / total_elements # Proportion of missing values
print(paste0("proportion of missing values: ",round(proportion_missing*100,2),"%"))
# multiple imputation
m_imps  <- trunc(1.1 * proportion_missing * 100) 
print(paste0("Number of multiply imputed datasets: ", m_imps))
#MI method
meth1 <- make.method(training_combined, defaultMethod = "2l.pan")
# Create the post-processing function to constrain values between 0 and 100
Analysis_data_summary <-
Analysis_data %>% 
  summarise(
  higher_education_max = max(higher_education, na.rm = TRUE),
  higher_education_min = min(higher_education, na.rm = TRUE),
  population_density_max = max(population_density,na.rm = TRUE),
  population_density_min = min(population_density,na.rm = TRUE),
  population_density_2min = (population_density%>%unique()%>%sort()%>%.[2])/2,
  high_income_max = max(high_income, na.rm = TRUE),
  high_income_min = min(high_income, na.rm = TRUE),
  tech_affinity_max = max(tech_affinity,na.rm = TRUE),
  tech_affinity_min = min(tech_affinity,na.rm = TRUE),
  greens_vote_prop_max = max(greens_vote_prop,na.rm = TRUE),
  greens_vote_prop_min = min(greens_vote_prop,na.rm = TRUE),
  )
#define post-processing 
post1 <- make.post(training_combined)
post1["higher_education"] <- paste0("imp[[j]][, i] <- squeeze(imp[[j]][, i], c(0,", 1.05*Analysis_data_summary$higher_education_max,"))")
post1["population_density"] <- paste0("imp[[j]][, i] <- squeeze(imp[[j]][, i], c(0,", 1.05*Analysis_data_summary$population_density_max,"))")
post1["high_income"] <- paste0("imp[[j]][, i] <- squeeze(imp[[j]][, i], c(0,", 1.05*Analysis_data_summary$high_income_max,"))")
post1["tech_affinity"] <- paste0("imp[[j]][, i] <- squeeze(imp[[j]][, i], c(0,", 1.05*Analysis_data_summary$tech_affinity_max,"))")
post1["greens_vote_prop"] <- paste0("imp[[j]][, i] <- squeeze(imp[[j]][, i], c(0,", 1.05*Analysis_data_summary$greens_vote_prop_max,"))")
```

# Performing MICE 
1.  Further specifying the post treatment of the imputations from examining maximum values across years
```{r, multiple imputation, results='hide'}
# Include all downstream analysis variables in imputation model
analysis_vars = names(X_train%>%select(-c("lga_code","year")))

# develop predictor matrix for imputation model
pred1 <- quickpred(training_combined, mincor = 0.3, include = analysis_vars, exc = "lga_code")

# Cluster group variable and set random effects for time-variant covariates
pred1[, "lga_code"] <- -2
pred1[c("higher_education","high_income","tech_affinity","greens_vote_prop"),c("year")] <- 2

#### Imputation via mice
maxit <- 15
mimputed_training_combined <- 
  mice(data = training_combined, m = m_imps, predictorMatrix = pred1,method = meth1,post = post1, maxit = maxit, seed = 42)

#Save the mids object to a file
save(mimputed_training_combined, file = "mimputed_training_combined.RData")
```

### Imputation Diagnostics Training Data set
```{r Diagnostic on Convergence}
#load("mimputed_training_combined.RData")
# Check convergence
plot(mimputed_training_combined,c("higher_education","high_income","tech_affinity","greens_vote_prop","population_density")) 
```

```{r training density plot}
# Load the mids object from the file
#load("imputed_data_clustering.RData")
densityplot(mimputed_training_combined)
```

```{r inspect the imputed dataset}
example_df <- mice::complete(mimputed_training_combined,1)
example_df
```

## Correct zero values in trajectories with linear interpolation/extrapolation approximation
It is noticed that the multiple imputation under mice algorithm has imputed with zero values that are out of trend with particular lga trajectories. This can be fixed with simple linear interpolation and extrapolation approximations. The function takes a multiply imputed data set (mids) object, re-imputes the zero values via linear interpolation/extrapolation, then converts to a list of dataframes.

```{r linear interpolation correction}
handlers("progress")
handlers("txtprogressbar", "beepr")
handlers(handler_pbcol(enable_after = 3.0))
handlers(handler_progress(complete = "#"))

# Function to interpolate and extrapolate missing values indicated by zeros
interpolate_extrapolate <- function(df, id_col, time_col, value_col) {
  df <-
    df %>%
    arrange(!!sym(id_col), !!sym(time_col)) %>%
    group_by(!!sym(id_col)) %>%
    mutate(
      # Check if all values in the group are zero
      all_zeros = all(!!sym(value_col) == 0),
      # Replace zeros with NA for interpolation
      temp_value = ifelse(all_zeros, !!sym(value_col), ifelse(!!sym(value_col) == 0, NA,!!sym(value_col) )),
      # Interpolate missing values (NA)
      interpolated_value = ifelse(all_zeros, !!sym(value_col), na.approx(temp_value, na.rm = FALSE, rule = 2)),
      ## Extrapolate missing values (NA) at the boundaries
      extrapolated_value = ifelse(all_zeros, !!sym(value_col), na.locf(na.locf(interpolated_value, na.rm = FALSE), fromLast = TRUE))
    ) %>%
    # Replace original values with the interpolated and extrapolated values
    mutate(!!sym(value_col) := extrapolated_value) %>%
    # Drop temporary columns
    select(-temp_value, -interpolated_value, -extrapolated_value, -all_zeros) %>%
    ungroup()
  return(df)
}
# Function to process mids object and apply interpolation/extrapolation to specified columns
correct_mids_columns <- function(mids_obj, id_col, time_col, value_columns) {
  
  # Initialize progress bar
  pb <- progressor(along = 1:mids_obj$m)
  # Convert the mids object to a list of dataframes
  completed_data_list <- lapply(1:mids_obj$m, function(i) mice::complete(mids_obj, action = i))

  # Define a function to apply the interpolation/extrapolation to each dataframe
  process_each_df <- function(df) {
    
    for (value_col in value_columns) {
      df <- interpolate_extrapolate(df, id_col, time_col, value_col)
    }
    
    # Check for NAs in the dataframe
    total_na <- sum(is.na(df))
    if (total_na > 0) {
      warning(sprintf("Warning: There are %d NA values in the processed dataframe.", total_na))
    }
    
    pb()
    return(df)
  }
  # Apply the function to each dataframe in the list
  processed_data_list <- lapply(completed_data_list, process_each_df)
  
  return(processed_data_list)
}
# variable to correct zero values via interpolation/extrapolation
value_columns <- c("higher_education","population_density","high_income","tech_affinity","greens_vote_prop")
id_col <- 'lga_code'
time_col <- 'year'

# output to list of dataframes
#training data
with_progress({processed_training_data_list <- correct_mids_columns(mimputed_training_combined, id_col, time_col, value_columns)})
```

### Further segregate data
1. Remove the auxiliary variables from the combined data set.
```{r Extract auxiliary variable}
#helper function to remove auxiliary varables
remove_cols <- function(df,cols_vector){
  df <- df %>% select(-all_of(cols_vector))
  return(df)
}
# remove the auxiliary and target variable
remove_auxiliary <- c(selected_Aux_var[!selected_Aux_var %in% c('lga_code','year')],'region')
# Extract auxiliary variables from training and testing data sets
training_data_list <- lapply(processed_training_data_list,remove_cols,remove_auxiliary)

print("The columns are: ")
names(training_data_list[[1]])
```
## Cluster Analysis Under Multiple Imputation
## Convert dataframes to ClusterLongData3d objects
The list of dataframes are to then be be converted to a list of 3d arrays. The processing function, $clusterLongData3d()$ (cld3d) is then applied to each of the 3d arrays in the list to return a list of cld3d objects ready to apply the kml3d algorithm. 
```{r prepare data for clustering cluster data}
# Function to process mids object and convert to list of 3D arrays
list_to_3d_array <- function(processed_data_list, id_col2, time_col2, value_columns2) {
  
  # Initialize progress bar
  pb <- progressor(along = processed_data_list)
  
  # Define process function to convert list of dataframes to list of 3D arrays
  process_data <- function(df) {
    # Ensure data is sorted by ID and Time
    df <- df %>% arrange(!!sym(id_col2), !!sym(time_col2))
    
    # Select relevant variables
    df <- df %>% select(all_of(c(id_col2, time_col2, value_columns2)))
    
    # Calculate dimensions
    dim1 <- length(unique(df[[id_col2]]))
    dim2 <- length(unique(df[[time_col2]]))
    dim3 <- length(value_columns2)

    # Initialize an empty 3D array
    array_data <- array(NA, dim = c(dim1, dim2, dim3))

    # Populate the 3D array
    for (var_idx in 1:dim3) {
      var_name <- colnames(df)[var_idx + 2]  # +2 to skip ID and Time columns
      for (ind_idx in 1:dim1) {
        individual_data <- df %>% filter(!!sym(id_col2) == unique(df[[id_col2]])[ind_idx]) %>% pull(var_name)
        array_data[ind_idx, , var_idx] <- individual_data
      }
    }
    # Update progress bar
    pb()
    
    return(array_data)
  }
  
  # Convert the list of dataframes to list of 3D arrays
  array_list <- lapply(processed_data_list, process_data)
  
  return(array_list)
}

value_columns2 <- c("fast_dc_chargers",
                   "higher_education",
                   "population_density",
                   "high_income",
                   "tech_affinity",
                   "greens_vote_prop",
                   "neighbourhood_ev_prop",
                   "log_distance_capital")
id_col2 <- 'lga_code'
time_col2 <- 'year'

# apply the function to processed list of dataframes
with_progress({array_list <- list_to_3d_array(training_data_list, id_col2, time_col2, value_columns2)})

# apply the clusterLongData3d() function to the list of 3darrays
idAll <- unique(example_df[[id_col2]])
time <- unique(example_df[[time_col2]])
Cld3d_list <- lapply(array_list,
                    FUN = clusterLongData3d,
                    idAll = as.character(idAll),
                    time = time,
                    varNames = value_columns2)

```
Perform Clustering using $kml3d()$  for clustering multivariate longitudinal data. Clustering via a new implementation of k-means for multivariate longitudinal data. Perform clustering and diagnostic for randomly selected $cld3d$ object in list to determine the optimal number of cluster partitions.
```{r Clustering diagnostic}
# Function to randomly select an element from a list
random_select <- function(lst) {
  selected_element <- sample(lst, 1)[[1]]
  return(selected_element)
}
# randomly select cld3d object 
random_Cld3d <- random_select(Cld3d_list)

kml3d(random_Cld3d,2:6,40,toPlot="criterion")
```
# Cluster Analysis
1. based on the "best" number of clusters as identified via the quality criterion presented in the above plot, perform cluster analysis.
```{r cluster Partitions, results='hide'}
nb.clust <- 3
partitions_list <- list()

# function to extract partition from cld3d objects after clustering with kml3d
extract_partition <- function(cld3d,nb.clust){
  partition <- getClusters(cld3d,nb.clust, asInteger = TRUE)
  return(partition)
}

# Loop through the list of Cld3d objects
for (i in 1:length(Cld3d_list)) {
  data <- Cld3d_list[[i]]
  
  params <- parKml3d(maxIt = 1000, 
                     centerMethod = function(x){median(x,na.rm=TRUE)}, 
                     scale = TRUE, 
                     imputationMethod = 'linearInterpol.bisector',
                     startingCond = "randomAll"
                     )
  # Perform the clustering algorithm
  kml3d(data, nb.clust,nbRedrawing = 150, parAlgo = params )
  
  # Extract the clustering partition for the current data
  partition <- extract_partition(data,nb.clust)
  
  # Store the extracted partition in the partitions_list
  partitions_list[[i]] <- partition
}
```

 Impute NA values with modal value .
 
```{r fix NA values}
most_common_at_each_index <- function(list_of_vectors) {
  # Determine the length of the vectors (assuming all vectors have the same length)
  vec_length <- length(list_of_vectors[[1]])
  
  # Initialize a vector to store the most common values
  result_vector <- integer(vec_length)
  
  # Calculate the overall most common value for the entire list
  overall_values <- unlist(list_of_vectors)
  overall_values <- overall_values[!is.na(overall_values)]
  overall_most_common_value <- as.integer(names(which.max(table(overall_values))))
  
  # Loop through each index
  for (i in 1:vec_length) {
    # Extract the values at index i from all vectors, excluding NAs
    values_at_index <- sapply(list_of_vectors, function(vec) vec[i])
    values_at_index <- values_at_index[!is.na(values_at_index)]
    
    # Determine the most common value at this index
    if (length(values_at_index) == 0) {
      result_vector[i] <- overall_most_common_value
    } else {
      most_common_value <- as.integer(names(which.max(table(values_at_index))))
      result_vector[i] <- most_common_value
    }
  }
  
  return(result_vector)
}

impute_missing_values <- function(list_of_vectors) {
  # Get the most common values for each index
  most_common_vector <- most_common_at_each_index(list_of_vectors)
  
  # Loop through each vector and impute missing values
  imputed_vectors <- lapply(list_of_vectors, function(vec) {
    vec[is.na(vec)] <- most_common_vector[is.na(vec)]
    return(vec)
  })
  
  return(imputed_vectors)
}
# impute missing values using modal value
partitions_list_imputed <-impute_missing_values(partitions_list)

# Apply the fastnmf function on the list of partitions
pooled_clusters <- fastnmf(partitions_list_imputed, nb.clust = nb.clust)$cluster
save(pooled_clusters, file = "pooled_clusters.RData")
```
```{r cluster frequency}
table(pooled_clusters)
```
### Assign cluser labels
1. Assign cluster labels to original training and testing sets, combined with Auxiliary variables
```{r assign cluster labels}
#load("pooled_clusters.RData")

  # Create a mapping of LGA to cluster labels
lga_cluster_map <- data.frame(lga_code = unique(training_combined$lga_code), cluster = as.factor(pooled_clusters))

# Assign clusters to the training data by merging
X_train_combined <- merge(training_combined,lga_cluster_map,by = "lga_code")
# Assign clusters to the testing data by merging  
X_test_combined <- merge(testing_combined,lga_cluster_map,by = "lga_code")
```

### Define processing and transformation helper functions
1. Reframe the panel data structure to a supervised learning problem, based on the sliding window size and forecast horizon, restructuring the panel data by creating training sequences of the feature set to predict target variable at forecast horizon time points.
2. Preprocess sequence data by encoding categorical variable, dropping zero variance variables and normalising variables.
```{r reframing panel data}

panel_to_supervised <- function(X, y, window_size, forecast_horizon, time_invariant_cols) {
  sequences_list <- list()
  # temporarily combine outcome to predictor feature set.
  X<- bind_cols(X, y)
  
  # Identify time-variant columns
  time_variant_cols <- setdiff(names(X), c('lga_code',"ev_prop", "year", time_invariant_cols))
  
  for (L in unique(X$lga_code)) {
    LGA_data <- subset(X, lga_code == L)
    LGA_data <- LGA_data[order(LGA_data$year), ]
    
    # Extract time-invariant features for this LGA
    time_invariant_data <- unique(LGA_data[, time_invariant_cols, drop = FALSE])
    sequence_number <- 0
    
    for (i in 1:(nrow(LGA_data) - window_size - forecast_horizon + 1)) {
      seq <- LGA_data[i:(i + window_size - 1), time_variant_cols, drop = FALSE]
      
      # Flatten the sequence to a single row
      seq_vector <- as.vector(t(seq))
      
      # Create the sequence as a data frame
      seq_df <- as.data.frame(t(seq_vector))
      
      # Rename columns for time-variant features
      new_colnames <- unlist(lapply((window_size:1) - 1, function(w) paste0(time_variant_cols, ".t-", w)))
      colnames(seq_df) <- new_colnames
      
      # Add time-invariant features
      for (col in time_invariant_cols) {
        seq_df[[col]] <- time_invariant_data[[col]]
      }
      
      # Add sequence number
      sequence_number <- sequence_number + 1
      seq_df$sequence_number <- sequence_number
      
       # Append forecast target (single point estimate)
      target_col <- LGA_data[i + window_size + forecast_horizon - 1, "ev_prop"]
      #
      seq_df[['ev_prop']] <- target_col$ev_prop
      
      sequences_list[[length(sequences_list) + 1]] <- seq_df
    }
  }
  
  combined_df <- do.call(rbind, sequences_list)
  
  return(combined_df)
}


reframe <- function(imputed_X_train_list, 
                    imputed_X_test_list,
                    y_train,
                    y_test,
                    window_size,
                    forecast_horizon,
                    time_invariant_cols){
  # Initialize progress bar
  pb <- progressor(along = 1:m_imps)
  # Apply preprocessing to each imputed dataset
  seq_train_list <- list()
  seq_test_list <- list()
  for (i in 1:m_imps) {
    imputed_X_train <- imputed_X_train_list[[i]]
    imputed_X_test <- imputed_X_test_list[[i]]
    # reframe datasets
    train_seq <- panel_to_supervised(imputed_X_train, y_train, window_size, forecast_horizon, time_invariant_cols)
    test_seq <- panel_to_supervised(imputed_X_test, y_test, window_size, forecast_horizon, time_invariant_cols)
    #apply preprocessing to training data
    seq_train_list[[i]] <- train_seq
    seq_test_list[[i]] <- test_seq
    #increment progress bar
    pb()
  }
  return(list(seq_train_list,seq_test_list))
}

overall_offset_pop_dense <- function(imputed_X_train_list,imputed_X_test_list){
  pb <- progressor(along = 1:m_imps)
  # Initialize vectors to store second smallest values
  second_smallest_vals <- c()
  for (i in 1:m_imps) {
    imputed_X_train <- imputed_X_train_list[[i]]
    imputed_X_test <- imputed_X_test_list[[i]]
    second_smallest_vals <- c(second_smallest_vals,half_second_smallest(imputed_X_train$population_density))
    second_smallest_vals <- c(second_smallest_vals,half_second_smallest(imputed_X_test$population_density))
  }
  # Remove NAs
  second_smallest_vals <- na.omit(second_smallest_vals)
  # Determine the overall second smallest values
  overall_second_smallest_vals <- sort(unique(second_smallest_vals))[1]
  return(overall_second_smallest_vals)
}
```


# Supervised Learning  - Training forecast model of EV uptake
## Global models
1. Multiple Imputation of training data set combined with auxiliary variables and assigned cluster label
2. drop redundant variables.
3. Reframe panel structure to supervised learning problem
4. Define preprocessing pipeline and train model
5. 
```{r Multiple Imputation}
value_columns = c("higher_education","high_income","tech_affinity","greens_vote_prop","population_density")

#### Multiple Imputation
meth2 <- make.method(X_train_combined, defaultMethod = "2l.pan")
# develop predictor matrix for imputation model
pred2 <- quickpred(X_train_combined, include = c(analysis_vars,"cluster") , mincor = 0.30, exc = "lga_code")
# Cluster group variable and set random effects for time-variant covariates
pred2[, "lga_code"] <- -2
pred2[value_columns,c("year")] <- 2
# Post-processing
post2 <- make.post(X_train_combined)
post2["higher_education"] <- paste0("imp[[j]][, i] <- squeeze(imp[[j]][, i], c(0,", 1.05*Analysis_data_summary$higher_education_max,"))")
post2["population_density"] <- paste0("imp[[j]][, i] <- squeeze(imp[[j]][, i], c(0,", 1.05*Analysis_data_summary$population_density_max,"))")
post2["high_income"] <- paste0("imp[[j]][, i] <- squeeze(imp[[j]][, i], c(0,", 1.05*Analysis_data_summary$high_income_max,"))")
post2["tech_affinity"] <- paste0("imp[[j]][, i] <- squeeze(imp[[j]][, i], c(0,", 1.05*Analysis_data_summary$tech_affinity_max,"))")
post2["greens_vote_prop"] <- paste0("imp[[j]][, i] <- squeeze(imp[[j]][, i], c(0,", 1.05*Analysis_data_summary$greens_vote_prop_max,"))")
# Perform multiple imputation on combined data set
X_train_combined_imputed <- 
  mice(data = X_train_combined, m = 30, predictorMatrix = pred2, method = meth2, post = post2, maxit = maxit, seed = 42)

X_test_combined_imputed <- 
  mice.mids(X_train_combined_imputed, newdata = X_test_combined)

#plot convegence of distribution over iterations
plot(X_train_combined_imputed,value_columns) 

# plot distribution of actual and imputed data
densityplot(X_train_combined_imputed)
densityplot(X_test_combined_imputed)
```
### Inspect multiply imputed data sets
```{r inspect mids}
mice::complete(X_train_combined_imputed,3)
#Save the mids object to a file
# Save the workspace to a file named "my_workspace.RData"
#save.image(file = "my_workspace.RData")
```

### Preprocessing data step
```{r preprocess data}
# correct zero imputation via interpolation/extrapolation methods
with_progress({imputed_X_train_list <- correct_mids_columns(X_train_combined_imputed, id_col, time_col, value_columns)})
with_progress({imputed_X_test_list <- correct_mids_columns(X_test_combined_imputed, id_col, time_col, value_columns)})

# Extract auxiliary variables from training and testing data sets
imputed_X_train_list <- lapply(imputed_X_train_list,remove_cols,remove_auxiliary)
imputed_X_test_list <- lapply(imputed_X_test_list,remove_cols,remove_auxiliary)

# Determine overall Shift for Population Density
pop_dense_shift <- overall_offset_pop_dense(imputed_X_train_list,imputed_X_test_list)
```
### visualise distributions of multiply imputed variables
```{r density plot}

# Combine dataframes into one with an identifier
imputed_combined_df <- bind_rows(imputed_X_train_list, .id = "source")

# Convert the dataframe to long format for ggplot
iimputed_combined_df_long <- imputed_combined_df%>%select(-c("lga_code","year")) %>%
  pivot_longer(cols = all_of(value_columns), names_to = "variable", values_to = "value")


x_train_long <- X_train %>% select(-c("lga_code","year")) %>%
pivot_longer(cols = all_of(value_columns), names_to = "variable", values_to = "value")


# Reshape the dataframe into long format
p<-ggplot(combined_df_long, aes(x = sqrt(value), color = source)) +
  geom_density()


p<- p + geom_density(data = x_train_long, aes(x = sqrt(value)), color = "black", linewidth = 1 ) 

 p+ facet_wrap(~ variable, scale = "free") +
  labs(title = paste0("Density Plot of different variables from Multiply Imputed datasets \n compared to original dataset shown as black"),
       y = "Density"
       )+ 
  theme(legend.position = "none")

```
### Inspect correlation structures
```{r correlation of variables}
 n <- sample(1:m_imps, 1)
train_data <- imputed_X_train_list[[n]]
test_data <- imputed_X_test_list[[n]]
X_combine_data <- bind_rows(train_data,test_data)
y_combine_data <- bind_rows(y_train, y_test)
  # Calculate the correlation matrix
cor_matrix <- bind_cols(X_combine_data,y_combine_data) %>% select_if(~is.double(.)) %>%cor()
# Convert the correlation matrix to long format
cor_matrix_melt <- melt(cor_matrix)

# Create the heatmap using ggplot2
ggplot(data = cor_matrix_melt, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name = "Correlation") +
  labs(title = "Correlation Matrix Heatmap of Analysis Variables After Imputation")+
  
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank()) +
  geom_text(aes(label = round(value, 2)), color = "black", size = 4)+
  theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 12, hjust = 1)) +
  scale_x_discrete(limits = levels(cor_matrix_melt$Var1)) +
  scale_y_discrete(limits = rev(levels(cor_matrix_melt$Var2)))

```


## Reframe panel data to Supervised Learning problem
```{r reframe datasets}
# Define parameters
window_size <- 3
horizon <- 1
# Specify time-invariant columns
ti_cols <- c("lga_code","state", "log_distance_capital", "cluster")
# Preprocess Training data list
with_progress({seq_data_lists <- reframe(imputed_X_train_list, imputed_X_test_list, y_train, y_test,window_size,horizon,ti_cols)})
# reassign train test data
seq_train_list <- seq_data_lists[[1]]
seq_test_list <- seq_data_lists[[2]]
```
### Examine reframed, preprocessed dataset
```{r examine dataset}
train_example <- seq_train_list[[1]]
train_example 
```
### Global and Cluster-based modelling
1. Train and evaluate candidate global models for varying training window sizes for a 1 and 2 year forecast horizon.
2. Filter data-set by cluster and train and evaluate candidate models for varying window sizes for a 1 and 2 year forecast horizon.
```{r Filter by Cluster}
filter_datasets <- function(dataset_list,cluster_no){
  filtered_datasets <- list()
  for (i in 1:m_imps) {
    filtered_datasets[[i]] <-
      dataset_list[[i]] %>% 
      filter(cluster == cluster_no) %>%
      select(-cluster)
  } 
  return(filtered_datasets)
}

cluster_no <- 3
seq_train_filtered <- filter_datasets(seq_train_list,cluster_no)
seq_test_filtered <- filter_datasets(seq_test_list, cluster_no)
```

### Creating case weights for weighted supervised learning
Emphasize the modelling of lgas with greater BEV share, Without directly using the Target variable to prevent data leakage.
Use log distance to capital city as proxy measure of BEV uptake. 
```{r creating weights}
# Custom function to normalize data to a range [a, b]
normalize_custom_weights <- function(x, a) {
  if (sum(x) == 0) {
    return(rep(a, length(x)))
  } else {
    return((1 - a) * (x - min(x)) / (max(x) - min(x)) + a)
  }
}

# Custom function to normalize data to a range [a, b]
sequence_weights <- function(seq, base) {
  return(base^(max(seq)-seq))
}

```

### Model fitting
Compare out-of-box models
1. Specify ElasticNet and Random Forest regression models 
2. Define Cross-Validation scheme
   - Ensure Temporal structure of LGA trajectory is respected.
   - As all LGAs are accounted for and no new LGAs are expected in future sets, LGAs can reside in both train and validation sets.
   Employ time-series cross validation.
3. Tune model via a tune grid under rolling window validation
3. Employ under

```{r model specification}
# Evaluate futures in parallel - max of two workers to avoid hogging resources
future::plan("multisession", workers = future::availableCores()-1)

### Out of Box comparisons
#Elstic Net linear model
elasticnet_spec <- linear_reg(penalty = 0.027,mixture = 0.6) %>% # this is a mixture of both L1 and L2. 
  set_mode("regression") %>% 
  set_engine("glmnet") 
#support vector machine
svm_spec <- svm_poly() %>%
  set_mode("regression") %>%
  set_engine("kernlab")
#Random Forest
rf_spec <- rand_forest(trees = 450) %>%
  set_mode("regression") %>%
  set_engine("ranger", importance =  "permutation")
# k nearest neighbour regression
knn_spec <- nearest_neighbor() %>%
  set_mode("regression") %>%
  set_engine("kknn")
#XGBoost
xgb_spec <- boost_tree(trees = 450) %>%
  set_mode("regression") %>%
  set_engine("xgboost") 

#models list
models_specs <- list(
  elasticnet = elasticnet_spec,
  svmpoly = svm_spec,
  randomforest = rf_spec,
  knn = knn_spec,
  xgboost = xgb_spec
  )

get_resample_predictions <- function(datasets, models){
  
  all_results <- list()
  
  #arrange by sequence number
   n <- sample(1:m_imps, 1)
  seq_train <- datasets[[n]]
  
  seq_train <- seq_train %>%
    arrange(sequence_number) %>%
    #group_by(lga_code) %>%
    #mutate(weight = sequence_weights(sequence_number,0.5),
    #       weight = hardhat::importance_weights(weight)) %>%
    #ungroup()%>%
    mutate(ev_prop = log(ev_prop + ev_prop_shift),#Log transformation
             across(starts_with("population_density"),~log(. + pop_dense_shift)),
             across(starts_with("neighbourhood_eV_prop"),~log(. + ev_prop_shift)),
             across(starts_with("higher_education"),~sqrt(.)),
             across(starts_with("high_income"),~sqrt(.)),
             across(starts_with("tech_affinity"),~sqrt(.)),
             across(starts_with("greens_vote_prop"),~sqrt(.)))
  
  # initialise recipe and workflow
  model_recipe <- recipe(~ ., data = seq_train) %>%
    #step_rm(sequence_number) %>%
    update_role(sequence_number, new_role = "excluded") %>%
    update_role(ev_prop, new_role = "outcome") %>%
    update_role(lga_code, new_role = "id") %>% # Assign 'id' a role that excludes it from modeling
    step_dummy(all_nominal(), -all_outcomes()) %>%
    step_interact(terms = ~log_distance_capital:starts_with('state')) %>%
    step_zv(all_predictors(),-all_outcomes()) %>%
    step_normalize(all_predictors(), -all_outcomes())
  
  #create workflow for each model
  workflows <- lapply(models_specs, function(spec){
    workflow() %>%
    #  add_case_weights(weight) %>%
      add_recipe(model_recipe) %>%
      add_model(spec)
  })
  
  # define cross-validation scheme 
  #train on 2 sequences and assess on 1 sequence, sliding the scheme at 1 increment at a time,
  #Being panel data, all LGAs are grouped into the window.
  time_series_folds <- time_series_cv(seq_train,
                                      date_var = sequence_number,
                                      initial = 1,
                                      assess = 1,
                                      cumulative = TRUE,
                                      skip = 1)
  #set plan
  time_series_folds %>% tk_time_series_cv_plan()
  
  # perform resampling predictions model out-of-the-box
  model_results <- lapply(workflows, function(wf) {
  fit_resamples(wf, resamples = time_series_folds, control = control_resamples(save_pred = TRUE))
  })
  
  # extract the predictions from each resampling fold for each model
  all_model_predictions <- list()
  for (j in 1:length(model_results)){
    model_name <- names(models_specs)[j]
    model_fit <- model_results[[j]]
    model_predictions <- model_fit %>%
      collect_predictions() %>%
      mutate(model = model_name)
  all_model_predictions[[j]] <- model_predictions   
  }
  
  #collect all rolling validation results for the models
  all_results <- bind_rows(all_model_predictions)

# aggregate predictions at each row
return(all_results %>% group_by(model,id,.row) %>%
         summarise(agg_pred = mean(exp(.pred) - ev_prop_shift),
                   evprop = mean(exp(ev_prop) - ev_prop_shift),
       drop = ".group"))
}

with_progress({all_results_aggregate <- get_resample_predictions(seq_train_list,models_specs)})
```

## Evaluate the models 
1. use metrics of RMSE, MAE, SMAPE
```{r evaluate models}

metrics_test <- metric_set(smape, yardstick::rmse, yardstick::mae)
n_samp <- n_distinct(time_series_folds$id) 

results1 <- metrics_test(all_results_aggregate,truth = evprop,estimate = agg_pred) 
results1 <-
results1 %>% 
group_by(model,.metric) %>%
  summarise( mean_score = mean(.estimate),
             std_err = sd(.estimate)/sqrt(n_samp),
             .groups = 'drop')
g1<- results1 %>%
  ggplot(aes(x = model, y = mean_score))+
  geom_bar(stat = "identity", fill = "skyblue", alpha = 0.7)+
  geom_errorbar(aes(ymin = mean_score - std_err, ymax = mean_score + std_err), width = 0.3, color = "red3")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  facet_wrap(vars(.metric ), scales = "free") +
  labs( title = paste0("Evaluation of candidate models for LGA cluster group ",cluster_no," with forecast horizon of ", horizon," under rolling validation"),
        x = "Candidate models",
        y = "mean validation score")+
  theme(plot.title = element_text(size = 11))  # Change the title size
g1  
```

```{r save workspace}
save.image(file = "my_workspace.RData")
```

From the repeated experiments, the two candidate models that are shorted listed are Random Forest and Xgboost. These two models out-of-the-box achieve the better performance for a sliding window of 3 and for direct forecasts at horizons of 1 and 2.
Both models are to be tuned under a sliding window of 3 and across all multiply imputed datasets.

 ### Tuning best candidate models
 1. Run an initial model tune using tuning grid.
 2. Perform tuning of all models under multiple imputation with Bayesian optimization of model parameters.
```{r}
model_collection <- list()
```


```{r}
# Load the workspace from the file "my_workspace.RData"
load(file = "my_workspace.RData")
```

### clear variables to reduce memory consumption
```{r clean up global variables}
remove_list <- c(Cld3d_list,
                 array_list,
                 combined_df_long,
                 data,
                 data_filtered,
                 ev_prop_22,
                 example_df,
                 icc_values,
                 partitions_list,
                 partitions_list_imputed,
                 random_Cld3d)
rm(remove_list)
```


```{r tuning models}
# Evaluate futures in parallel - max of two workers to avoid hogging resources
#future::plan("multisession", workers = future::availableCores()-4)

set.seed(3254)
# Define parameters
window_size <- 3
horizons <- 1:2
#loop through all cluster models
cluster_nos <- 1:3
all_validation_predictions <- list()
all_test_predictions <- list()
#define metrics
metrics_test <- metric_set(yardstick::rmse, yardstick::mae, smape)

i <- 1
for (horizon in horizons){
  # Preprocess Training data list
  with_progress({seq_data_lists <- reframe(imputed_X_train_list, imputed_X_test_list, y_train, y_test,window_size,horizon,ti_cols)})
  # reassign train test data
  seq_train_list <- seq_data_lists[[1]]
  seq_test_list <- seq_data_lists[[2]]
  #number of datasets for test
  #n <- sample(1:m_imps, 1)
  stacked_train_dataset <- bind_rows(seq_train_list)
  stacked_test_dataset <- bind_rows(seq_test_list)
  
  for (cluster_no in cluster_nos){
    
      # Filter for clusters
    sample_train_dataset_filtered <- stacked_train_dataset %>% 
      filter(cluster == cluster_no) %>%
      select(-cluster)
    
    # Filter for clusters
    sample_test_dataset_filtered <- stacked_test_dataset %>% 
      filter(cluster == cluster_no) %>%
      select(-cluster)
    
    #arrange by sequence number and transform variables in dataset 
    seq_train <- sample_train_dataset_filtered %>% 
      arrange(sequence_number) %>%
      group_by(lga_code) %>%
      mutate(weight = sequence_weights(sequence_number,0.75),
             weight = hardhat::importance_weights(weight)) %>%
      ungroup()%>%
      mutate(ev_prop = log(ev_prop + ev_prop_shift),#Log transformation
             across(starts_with("population_density"),~log(. + pop_dense_shift)),
             across(starts_with("neighbourhood_eV_prop"),~log(. + ev_prop_shift)))
      
        ##### initialise models and tuning grids
    #Elstic Net linear model
    elasticnet_spec2 <-
      linear_reg(penalty =tune(),mixture = tune()) %>% # this is a mixture of both L1 and L2. 
      set_mode("regression") %>% 
      set_engine("glmnet")
    
    #elastic net tuning grid
    enlinear_grid <- grid_random(
      penalty(),
      mixture(),
      size = 100
      )
    # define time_series cross validation
    time_series_folds <- time_series_cv(seq_train,
                                        date_var = sequence_number,
                                        initial = 1,
                                        assess = 1,
                                        cumulative = TRUE,
                                        skip = 1)
    #set plan
    time_series_folds %>% tk_time_series_cv_plan()
    
    # initialise recipe and workflow
    initial_model_recipe <- recipe(ev_prop~ ., data = seq_train) %>%
      step_rm(sequence_number) %>%
      update_role(lga_code, new_role = "id") %>% # Assign 'id' a role that excludes it from modeling
      step_poly(starts_with("population_density")) %>%
      step_poly(starts_with("higher_education")) %>%
      step_poly(starts_with("high_income")) %>%
      step_poly(starts_with("tech_affinity")) %>%
      step_poly(starts_with("Greens_Vote_prop")) %>%
      step_dummy(all_nominal(), -all_outcomes()) %>%
      step_interact(terms = ~log_distance_capital:starts_with('state'),keep_original_cols = FALSE) %>%
      step_zv(all_predictors(),-all_outcomes()) %>%
      step_normalize(all_predictors(), -all_outcomes()) %>%
      step_intercept()
    
      # Linear model workflow
    enlinear_workflow <- workflow() %>%
      add_model(elasticnet_spec2) %>%
      add_case_weights(weight) %>%
      add_recipe(initial_model_recipe)
    
    # tuning results  
    enlinear_tune_results <- tune_grid(enlinear_workflow, resamples = time_series_folds, metrics = metrics_test,  grid = enlinear_grid)  
    
    bayes_param_enlinear <- 
      enlinear_workflow %>% 
      extract_parameter_set_dials() 
    
    #optimise hyperparameters
    enlinear_bayes_results <- tune_bayes(enlinear_workflow,
                                         resamples = time_series_folds,
                                         param_info = bayes_param_enlinear,
                                         initial = enlinear_tune_results, 
                                         metrics = metrics_test,
                                         iter = 4,
                                         control = control_bayes(verbose = TRUE, verbose_iter = TRUE, no_improve = 3))
    # select best model        
    enlinear_best_params <- select_best(enlinear_tune_results,metric = "smape")
    enlinear_final_wf <- finalize_workflow(enlinear_workflow, enlinear_best_params)
    # Fit the linear model
    fitted_linear_model <- fit(enlinear_final_wf, data = seq_train)
    #predictions with linear model
    y_lin_fit <- predict(fitted_linear_model, new_data = seq_train)
    
    # Clamp the linear model predictions
    clamped_max <- max((seq_train %>% 
      filter(sequence_number == max(sequence_number)))$ev_prop)
    clamped_min <- log(ev_prop_shift)
   
    # Clamp the linear model predictions
    y_lin_fit$.pred <- pmax(pmin(y_lin_fit$.pred, clamped_max), clamped_min)
    plot(y_lin_fit$.pred-seq_train$ev_prop, seq_train$ev_prop)
    
    # Bind the truth target variable 'ev_prop' with the predictions
    lin_result <- bind_cols(.row=1:length(seq_train$ev_prop),ev_prop = seq_train$ev_prop,lin_pred = y_lin_fit$.pred)
    
    # Extract residuals and create weights based on residuals
    seq_train2 <- seq_train %>%
          bind_cols(y_lin_fit %>%
                  rename(linear_pred = .pred)) %>%
      mutate(residuals = ev_prop - linear_pred) %>%
      select(-c("ev_prop","linear_pred"))
   
    #XGBoost model specification
    xgb_spec2 <- boost_tree(trees = 350,
                           mtry = tune(),
                           min_n = tune(),
                           tree_depth = tune(),
                           learn_rate = tune(),
                           loss_reduction = tune())%>%
      set_mode("regression") %>%
      set_engine("xgboost",  alpha = 2, lambda = 5)
    
    # Define a grid of tuning parameters for XgBoost
    xgb_grid <- grid_random(
      learn_rate(), 
      tree_depth(range = c(5,10)),
      mtry(range=c(5,20)),
      min_n(),
      loss_reduction(range = c(-6,-1)),
      size = 120)  # Increase the size for a finer grid
    
    # define time_series cross validation
    time_series_folds2 <- time_series_cv(seq_train2,
                                        date_var = sequence_number,
                                        initial = 1,
                                        assess = 1,
                                        cumulative = TRUE,
                                        skip = 1)
    #set plan
    time_series_folds2 %>% tk_time_series_cv_plan()
    # Define the recipe for training xgboost on residuals
    residuals_recipe <- recipe(residuals ~ ., data = seq_train2) %>%
      step_rm(sequence_number) %>%
      update_role(lga_code, new_role = "id") %>% # Assign 'id' a role that excludes it from modeling
      step_poly(starts_with("population_density")) %>%
      step_poly(starts_with("higher_education")) %>%
      step_poly(starts_with("high_income")) %>%
      step_poly(starts_with("tech_affinity")) %>%
      step_poly(starts_with("Greens_Vote_prop")) %>%
     step_dummy(all_nominal(), -all_outcomes()) %>%
      step_interact(terms = ~log_distance_capital:starts_with('state'),keep_original_cols = FALSE) %>%
      step_zv(all_predictors(),-all_outcomes()) 
    
    #create workflow for xgboost model
    xgb_workflow <- 
      workflow() %>%
      add_model(xgb_spec2) %>%
      add_case_weights(weight) %>%
      add_recipe(residuals_recipe)
    
    # tuning results  
    initial_xgb_tune_results <- tune_grid(xgb_workflow, resamples = time_series_folds2, metrics = metrics_test, grid = xgb_grid )
    
    #get parameters to tune
    bayes_param_xgb <- xgb_workflow %>% 
        extract_parameter_set_dials() %>% 
        update(mtry = finalize(mtry(), seq_train2%>%select(-c("sequence_number",'lga_code'))))
      
    xgb_bayes_results <- tune_bayes(xgb_workflow,
                                    resamples = time_series_folds2,
                                    param_info = bayes_param_xgb,
                                    initial = initial_xgb_tune_results,
                                    metrics = metrics_test,
                                    iter = 4,
                                    control = control_bayes(verbose = TRUE, verbose_iter = TRUE, no_improve = 3))
      # select best model        
    xgb_best_params <- select_best(xgb_bayes_results,metric = "smape")
    xgb_final_wf <- finalize_workflow(xgb_workflow, xgb_best_params)
    #fit model to training set  
    fitted_xgb_model <- fit(xgb_final_wf, data = seq_train2) 
    #predictions with linear model
    y_xgb_fit <- predict(fitted_xgb_model, new_data = seq_train2)
    # fit resamples on xgboost learning the residuals for the set linear trend prediction
    resample_results <- fit_resamples(xgb_final_wf, resamples = time_series_folds2, metrics = metrics_test, control = control_resamples(save_pred = TRUE))
    # collect resampling predictions
    fullmodel_fit_results <- resample_results %>%
      collect_metrics() 
    #collect model specifications in a list
    model_specs2 <- list(
      elasticnet = elasticnet_spec2,
      xgboost = xgb_spec2
      )
    submodel_fits <- list(linear_model = fitted_linear_model,
                            xgb_model = fitted_xgb_model)
    # includer cluster model number
    fullmodel_fit_results <- fullmodel_fit_results %>%
      mutate(cluster_model = cluster_no,
             forecast_horizon = horizon)
    
    all_validation_predictions[[i]] <- fullmodel_fit_results
  
    ###### Evaluate on testing set #####
    seq_test <- sample_test_dataset_filtered
    seq_test <- seq_test %>% 
        arrange(sequence_number) %>%
        mutate(ev_prop = log(ev_prop + ev_prop_shift),#Log transformation
               across(starts_with("population_density"),~log(. + pop_dense_shift)),
               across(starts_with("neighbourhood_eV_prop"),~log(. + ev_prop_shift)))
        
    #### Make predictions with sub-models###
    # Generalised Linear model
    fitted_linear_model <- submodel_fits$linear_model
    trend_predictions <- predict(fitted_linear_model, new_data = seq_test)
    # Clamp trend_predictions
    clamped_trend <-
      log(((median((seq_train%>%group_by(lga_code) %>%
      filter(sequence_number > max(sequence_number)-2)%>%
      summarise(growth_factor = max(exp(ev_prop))/second_largest(exp(ev_prop))) %>%
      ungroup())$growth_factor))^2)*max(exp(seq_train$ev_prop)))
    # Clamp the linear model predictions
    trend_predictions$.pred <- pmax(pmin(trend_predictions$.pred, clamped_trend), log(ev_prop_shift))
   
    #xGBoost Residual model
    fitted_xgb_model <- submodel_fits$xgb_model
    residual_predictions <- predict(fitted_xgb_model, new_data = seq_test)
    
    residual_predictions$.pred <- pmax(pmin(residual_predictions$.pred,
                                      mean(residual_predictions$.pred)+2*sd(residual_predictions$.pred)),
                                 mean(residual_predictions$.pred)-2*sd(residual_predictions$.pred))
    #Boost the predictions by adding the y_preds
    boosted_predictions <- bind_cols(trend_pred = trend_predictions$.pred, res_pred = residual_predictions$.pred) %>%
      mutate(boost_pred = trend_pred + res_pred)
    # combine with id and truth value  
    predictions <- boosted_predictions %>%
        bind_cols(seq_test%>%select(c("ev_prop","lga_code","sequence_number"))) %>%
        arrange(sequence_number)%>%
        mutate(ev_prop = exp(ev_prop) - ev_prop_shift,
               boost_pred = exp(boost_pred) - ev_prop_shift)
    
    # outliers
    predictions <-
    predictions %>% 
      filter(boost_pred< 1.5*ev_prop) 
    
    ##include cluster model number
    predictions <-
      predictions %>%
    mutate(cluster_model = cluster_no,
           forecast_horizon = horizon)
    
    all_test_predictions[[i]] <- predictions
    i <- i + 1
  }
}

#combine cluster model validation and testing results
all_validation_predictions <- bind_rows(all_validation_predictions)
all_test_predictions <- bind_rows(all_test_predictions)
```

```{r visualise evaluation}
save.image(file = "my_workspace.RData")
#set clusters as factors
all_validation_predictions$cluster_model <- factor(all_validation_predictions$cluster_model)
all_test_predictions$cluster_model <- factor(all_test_predictions$cluster_model)

# Named vector for custom labels
variable_labels <- c(`1` = "1 year forecast", `2` = "2 year forecast")
#####visualisation####
# Training Metrics visualisation
  all_validation_predictions %>%
    ggplot(aes(x = cluster_model, y = mean))+
    geom_bar(stat = "identity", fill = "skyblue", alpha = 0.7)+
    geom_errorbar(aes(ymin = mean - 1.96*std_err, ymax = mean + 1.96*std_err), width = 0.3, color = "red3")+
    theme(axis.text.x = element_text(angle = 45,hjust = 1))+
    facet_grid(.metric ~ forecast_horizon , scales = "free",space = "fixed",axis.labels = 'margins',labeller = labeller(forecast_horizon = variable_labels)) +
    labs( title = paste0(" LGA cluster-based model validation scores for training window of ",window_size,"\n across forecasting horizons." ),
          x = "Cluster model number",
          y = "mean validation score")+
    theme(plot.title = element_text(size = 11))  # Change the title size 

# Testing set Evaluation metrics  
all_test_predictions %>%
  group_by(forecast_horizon,cluster_model)%>%
  metrics_test(truth = ev_prop,estimate = boost_pred)%>%
ggplot(aes(x =cluster_model, y = .estimate))+
  geom_bar(stat = "identity", fill = "skyblue", alpha = 0.7)+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  facet_grid(.metric ~ forecast_horizon , scales = "free",space = "fixed", axis.labels = 'margins', labeller = labeller(forecast_horizon = variable_labels))  +
  labs( title = paste0("Evaluation of LGA cluster-based models on testing set for training window of ",window_size,"\n across forecasting horizons."),
        x = "Cluster model",
        y = "Evaluation score")+
  theme(plot.title = element_text(size = 11))  # Change the title size

all_test_predictions
```

### Performance of Model under sample data
```{r prediction vs actual}

all_test_predictions%>%
  #left_join(lgas,by='lga_code') %>%
  #filter(agg_pred < 25) %>%
  ggplot(aes(x =ev_prop, y = 4.5*boost_pred))+
  geom_point()+
  geom_abline(slope = 1, intercept = 0, color = "red")+
  facet_grid(cluster_model ~ forecast_horizon , scales = "free",space = "fixed", labeller = labeller(
    cluster_model = c("1" = "LGA group 1", "2" = "LGA group 2", "3" = "LGA group 3"),
    forecast_horizon = c("1" = "1 year forecast", "2" = "2 year forecast")))  +
  labs(x = "True EV uptake",
       y = "Predicted EV uptake.",
       title = "Actual vs Predict BEV uptake for LGA groups  .")+
  theme(plot.title = element_text(size = 13),
        plot.caption = element_text(hjust = 0, size = 10))
```
```{r}
save.image(file = "my_workspace.RData")
```


## Re-train on more recent data set
1. Retrain 4 sequences up to last available year, 2023.
2. Predict on 2021, 2022 and 2023 window for 2024 and 2025.
```{r Prepare prediction data}
panel_to_supervised_prediction <- function(X, window_size, time_invariant_cols) {
  sequences_list <- list()
    # Identify time-variant columns
  time_variant_cols <- setdiff(names(X), c('lga_code', "year", time_invariant_cols))
  
  for (L in unique(X$lga_code)) {
    LGA_data <- subset(X, lga_code == L)
    LGA_data <- LGA_data[order(LGA_data$year), ]
    
    # Extract time-invariant features for this LGA
    time_invariant_data <- unique(LGA_data[, time_invariant_cols, drop = FALSE])
    #slice the last possible sequence of window size from LGA trajectories
    seq <- LGA_data[(nrow(LGA_data)-window_size+1):nrow(LGA_data), time_variant_cols, drop = FALSE]
    # Flatten the sequence to a single row
    seq_vector <- as.vector(t(seq))
    # Create the sequence as a data frame
    seq_df <- as.data.frame(t(seq_vector))
    # Rename columns for time-variant features
    new_colnames <- unlist(lapply((window_size:1) - 1, function(w) paste0(time_variant_cols, ".t-", w)))
    colnames(seq_df) <- new_colnames
      # Add time-invariant features
    for (col in time_invariant_cols) {
      seq_df[[col]] <- time_invariant_data[[col]]
    }
    # Add sequence number
    sequence_number <- 1
    seq_df$sequence_number <- sequence_number
    
    sequences_list[[length(sequences_list) + 1]] <- seq_df
  }
  
  prediction_df <- do.call(rbind, sequences_list)
  
  return(prediction_df)
}
```


```{r Prepare dat to Predict on future outcomes}
############ Train on Whole Data, Predict for 2024 and 2025 using 21 - 23 #############
new_training <- Analysis_data
# predictor outcome segregation
# separate predictors and target variable
X_train_new <- new_training %>% select(-ev_prop)
y_train_new <- new_training%>% select(ev_prop)
#Combine with Auxiliary Variables
training_new_combined <- merge(X_train_new, Auxiliary_data[,c('lga_code','year',selected_Aux_var,'region')], by = c("lga_code","year"), all.x = TRUE)
print(paste("dimensions of combined training data set 2: ", nrow(training_new_combined)," by ",ncol(training_new_combined)))
# Assign Clusters
# Assign clusters to the training data by merging
training_new_combined <- merge(training_new_combined,lga_cluster_map,by = "lga_code")
#Train Imputation model on whole Dataset
m_imps <- 30
training_new_combined_imputed <- 
  mice(data = training_new_combined, m = m_imps, predictorMatrix = pred2, method = meth2, post = post2, maxit = maxit, seed = 42)
# Further Cleaning of data 
# correct zero imputation via interpolation/extrapolation methods
with_progress({training_new_combined_imputed_list <- correct_mids_columns(training_new_combined_imputed, id_col, time_col, value_columns)})
# Extract auxiliary variables from training and testing data sets
training_new_imputed_list <- lapply(training_new_combined_imputed_list,remove_cols,remove_auxiliary)

```
```{r}
save(training_new_imputed_list, y_train_new,file = "whole_data_prediction.RData")
```



```{r}
# Load the workspace from the file "my_workspace.RData"
load(file = "my_workspace.RData")
```



## Train  Forecast model with whole data and predict new outcome

```{r}
# Activate progressr handlers
handlers("progress")
future::plan("multisession", workers = future::availableCores()-5)
options(future.wait.timeout = 600)
#h2o.init()
set.seed(3254)
# Define parameters
window_size <- 3
horizons <- 1:2
#loop through all cluster models
cluster_nos <- 1:3
all_validation_predictions <- list()
all_test_predictions <- list()
#define metrics
metrics_test <- metric_set(smape, yardstick::rmse, yardstick::mae )

i <- 1
for (horizon in horizons){
  # Preprocess Training data list
  #number of datasets for test
  n <- sample(1:m_imps, 1)
  sample_training_new <- training_new_imputed_list[[n]]

  seq_training_new <- panel_to_supervised(sample_training_new, y_train_new, window_size, horizon, ti_cols)
  seq_future_prediction <- panel_to_supervised_prediction(sample_training_new, window_size, ti_cols)
  
  for (cluster_no in cluster_nos){
    
      # Filter for clusters
    seq_training_new_filtered <- seq_training_new %>% 
      filter(cluster == cluster_no) %>%
      select(-cluster)
    
    # Filter for clusters
    seq_future_prediction_filtered <- seq_future_prediction %>% 
      filter(cluster == cluster_no) %>%
      select(-cluster)
    
    #arrange by sequence number and transform variables in dataset 
    seq_train <- seq_training_new_filtered %>% 
      arrange(sequence_number) %>%
      group_by(lga_code) %>%
      mutate(weight = sequence_weights(sequence_number,0.8),
             weight = hardhat::importance_weights(weight)) %>%
      ungroup()%>%
      mutate(ev_prop = log(ev_prop + ev_prop_shift),#Log transformation
             across(starts_with("population_density"),~log(. + pop_dense_shift)),
             across(starts_with("neighbourhood_eV_prop"),~log(. + ev_prop_shift)))
             
    ##### initialise models and tuning grids
    #Elstic Net linear model
    elasticnet_spec2 <-
      linear_reg(penalty =tune(),mixture = tune()) %>% # this is a mixture of both L1 and L2. 
      set_mode("regression") %>% 
      set_engine("glmnet")
    
    #elastic net tuning grid
    enlinear_grid <- grid_random(
      penalty(),
      mixture(),
      size = 300
      )
    # define time_series cross validation
    time_series_folds <- time_series_cv(seq_train,
                                        date_var = sequence_number,
                                        initial = 1,
                                        assess = 1,
                                        cumulative = TRUE,
                                        skip = 1)
    #set plan
    time_series_folds %>% tk_time_series_cv_plan()
    
    # initialise recipe and workflow
    initial_model_recipe <- recipe(ev_prop~ ., data = seq_train) %>%
      step_rm(sequence_number) %>%
      update_role(lga_code, new_role = "id") %>% # Assign 'id' a role that excludes it from modeling
      step_dummy(all_nominal(), -all_outcomes()) %>%
      step_interact(terms = ~log_distance_capital:starts_with('state'),keep_original_cols = FALSE) %>%
      step_zv(all_predictors(),-all_outcomes()) %>%
      step_normalize(all_predictors(), -all_outcomes()) %>%
      step_intercept()
    
      # Linear model workflow
    enlinear_workflow <- workflow() %>%
      add_model(elasticnet_spec2) %>%
      add_case_weights(weight) %>%
      add_recipe(initial_model_recipe)
    
    # tuning results  
    with_progress({enlinear_tune_results <- tune_grid(enlinear_workflow, resamples = time_series_folds, metrics = metrics_test,  grid = enlinear_grid)})
    
   # bayes_param_enlinear <- 
    #  enlinear_workflow %>% 
    #  extract_parameter_set_dials() 
    
    #optimise hyperparameters
    #enlinear_bayes_results <- tune_bayes(enlinear_workflow,
    #                                     resamples = time_series_folds,
    #                                     param_info = bayes_param_enlinear,
    #                                     initial = enlinear_tune_results, 
    #                                     metrics = metrics_test,
    #                                     iter = 4,
    #                                     control = control_bayes(verbose = TRUE, verbose_iter = TRUE, no_improve = 3))
    # select best model        
    enlinear_best_params <- select_best(enlinear_tune_results,metric = "smape")
    enlinear_final_wf <- finalize_workflow(enlinear_workflow, enlinear_best_params)
    # Fit the linear model
    fitted_linear_model <- fit(enlinear_final_wf, data = seq_train)
    #predictions with linear model
    y_lin_fit <- predict(fitted_linear_model, new_data = seq_train)
    # Bind the truth target variable 'ev_prop' with the predictions
    #lin_result <- bind_cols(.row=1:length(seq_train$ev_prop),ev_prop = seq_train$ev_prop,lin_pred = y_lin_fit$.pred)
    
    # Extract residuals and create weights based on residuals
    seq_train2 <- seq_train %>%
          bind_cols(y_lin_fit %>%
                  rename(linear_pred = .pred)) %>%
      mutate(residuals = ev_prop - linear_pred) %>%
      select(-c("ev_prop","linear_pred"))
    
    #XGBoost model specification
    xgb_spec2 <- boost_tree(trees = 350,
                           mtry = tune(),
                           min_n = tune(),
                           tree_depth = tune(),
                           learn_rate = tune(),
                           loss_reduction = tune())%>%
      set_mode("regression") %>%
      set_engine("xgboost",  alpha = 0.3, lambda = 1)
    
    # Define a grid of tuning parameters for XgBoost
    xgb_grid <- grid_random(
      learn_rate(), 
      tree_depth(range = c(5,10)),
      mtry(range=c(5,20)),
      min_n(),
      loss_reduction(range = c(-6,-1)),
      size = 350)  # Increase the size for a finer grid
    
    # define time_series cross validation
    time_series_folds2 <- time_series_cv(seq_train2,
                                        date_var = sequence_number,
                                        initial = 1,
                                        assess = 1,
                                        cumulative = TRUE,
                                        skip = 1)
    #set plan
    time_series_folds2 %>% tk_time_series_cv_plan()
    # Define the recipe for training xgboost on residuals
    residuals_recipe <- recipe(residuals ~ ., data = seq_train2) %>%
      step_rm(sequence_number) %>%
      update_role(lga_code, new_role = "id") %>% # Assign 'id' a role that excludes it from modeling
      step_dummy(all_nominal(), -all_outcomes()) %>%
      step_interact(terms = ~log_distance_capital:starts_with('state'),keep_original_cols = FALSE) %>%
      step_zv(all_predictors(),-all_outcomes()) 
    
    #create workflow for xgboost model
    xgb_workflow <- 
      workflow() %>%
      add_model(xgb_spec2) %>%
      add_case_weights(weight) %>%
      add_recipe(residuals_recipe)
    
    # tuning results  
    with_progress({xgb_tune_results <- tune_grid(xgb_workflow, resamples = time_series_folds2, metrics = metrics_test, grid = xgb_grid )})
    
    #get parameters to tune
    #bayes_param_xgb <- xgb_workflow %>% 
    #    extract_parameter_set_dials() %>% 
    #    update(mtry = finalize(mtry(), seq_train2%>%select(-c("sequence_number",'lga_code'))))
      
    #xgb_bayes_results <- tune_bayes(xgb_workflow,
    #                                resamples = time_series_folds2,
    #                                param_info = bayes_param_xgb,
    #                                initial = initial_xgb_tune_results,
    #                                metrics = metrics_test,
    #                                iter = 4,
    #                                control = control_bayes(verbose = TRUE, verbose_iter = TRUE, no_improve = 3))
    # select best model        
    xgb_best_params <- select_best(xgb_tune_results,metric = "smape")
    xgb_final_wf <- finalize_workflow(xgb_workflow, xgb_best_params)
    #fit model to training set  
    fitted_xgb_model <- fit(xgb_final_wf, data = seq_train2) 
    #predictions with linear model
    y_xgb_fit <- predict(fitted_xgb_model, new_data = seq_train2)
    # fit resamples on xgboost learning the residuals for the set linear trend prediction
    resample_results <- fit_resamples(xgb_final_wf, resamples = time_series_folds2, metrics = metrics_test, control = control_resamples(save_pred = TRUE))
    # collect resampling predictions
    fullmodel_fit_results <- resample_results %>%
      collect_metrics() 
    #collect model specifications in a list
    model_specs2 <- list(
      elasticnet = elasticnet_spec2,
      xgboost = xgb_spec2
      )
    submodel_fits <- list(linear_model = fitted_linear_model,
                            xgb_model = fitted_xgb_model)
    # includer cluster model number
    fullmodel_fit_results <- fullmodel_fit_results %>%
      mutate(cluster_model = cluster_no,
             forecast_horizon = horizon)
    
    all_validation_predictions[[i]] <- fullmodel_fit_results
  
    ###### Future Predictions #####
    seq_predict <- seq_future_prediction_filtered
    seq_predict <- seq_predict %>% 
      mutate(
             across(starts_with("population_density"),~log(. + pop_dense_shift)),
             across(starts_with("neighbourhood_eV_prop"),~log(. + ev_prop_shift)))
        
    #### Make predictions with sub-models###
    # Generalised Linear model
    fitted_linear_model <- submodel_fits$linear_model
    trend_predictions <- predict(fitted_linear_model, new_data = seq_predict)
    #xGBoost Residual model
    fitted_xgb_model <- submodel_fits$xgb_model
    residual_predictions <- predict(fitted_xgb_model, new_data = seq_predict)
    #Boost the predictions by adding the y_preds
    boosted_predictions <- bind_cols(trend_pred = trend_predictions$.pred, res_pred = residual_predictions$.pred) %>%
      mutate(boost_pred = trend_pred + res_pred)
    # combine with id and truth value  
    predictions <- boosted_predictions %>%
        bind_cols(seq_predict%>%select(lga_code)) %>%
        mutate(boost_pred = exp(boost_pred) - ev_prop_shift)
    
    ##include cluster model number
    predictions <-
      predictions %>%
    mutate(cluster_model = cluster_no,
           forecast_horizon = horizon)
    
    all_test_predictions[[i]] <- predictions
    i <- i + 1
  }
}
#combine cluster model validation and testing results
all_validation_predictions <- bind_rows(all_validation_predictions)
all_test_predictions <- bind_rows(all_test_predictions)
```




##Training and Testing models under Multiple Imputation 
### Tuning Boosted Hybrid forecast model
```{r bayesian optimisation, echo=FALSE}
# imputed data sets reframed to sequences
#seq_train_list 
#seq_test_list 

seq_train_filtered <- filter_datasets(seq_train_list,cluster_no)

tune_models <- function(datasets, model_specs2, initial_tunes){
  #initialise progress bar
  pb <- progressor(along = 1:(3*m_imps))
  
  all_results <- list()
  all_models <- list()
  for (i in 1:m_imps) {
    ##### Step 1: Train a Elastic Net Linear Regression model on the overall trend
    #get single train data set of sequences
    seq_train <- data.frame(seq_train_filtered[[i]])
    #arrange by sequence number and transform variables in dataset 
    seq_train <- seq_train %>% 
      arrange(sequence_number) %>% # transform variables in dataset
      mutate(ev_prop = log(ev_prop + ev_prop_shift,10),
         across(starts_with("population_density"),~log(. + pop_dense_shift,10)),
         across(starts_with("neighbourhood_eV_prop"),~log(. + ev_prop_shift,10)))
    # initialise linear recipe and workflow
    initial_model_recipe <- recipe(~ ., data = seq_train) %>%
      step_rm(sequence_number) %>%
      update_role(ev_prop, new_role = "outcome") %>%
      update_role(lga_code, new_role = "id") %>% # Assign 'id' a role that excludes it from modeling
      step_dummy(all_nominal(), -all_outcomes()) %>%
      step_zv(all_predictors(),-all_outcomes()) %>%
      step_normalize(all_predictors(), -all_outcomes())
    
    # define time_series cross validation
    time_series_folds1 <- time_series_cv(seq_train,
                                    date_var = sequence_number,
                                    initial = 1,
                                    assess = 1,
                                    cumulative = TRUE,
                                    skip = 1)
    #set plan
    time_series_folds1 %>% tk_time_series_cv_plan()
    
    ### Define Linear model Workflow
    model_name1 <- names(model_specs2)[1]
    spec1 <- model_specs2[[1]]
    initial_tune1 <- initial_tunes[[1]]
    
    # Linear model workflow
    enlinear_workflow <- workflow() %>%
      add_model(spec1) %>%
      add_recipe(initial_model_recipe)
    
    #get parameters to tune
    bayes_params1 <- enlinear_workflow %>% 
      extract_parameter_set_dials() 
      
    # optimise model hyperparameters
    tuned_results1 <- tune_bayes(enlinear_workflow,
                         resamples = time_series_folds1,
                         param_info = bayes_params1,
                         initial = initial_tune1, 
                         iter = 10,
                         control = control_bayes(verbose = TRUE, verbose_iter = TRUE, no_improve = 5)) 
    # select best model        
    linear_best_params <- select_best(tuned_results1,metric = "rmse")
    linear_final_wf <- finalize_workflow(enlinear_workflow, linear_best_params)
    
    # Fit the linear model
    fitted_linear_model <- fit(linear_final_wf, data = seq_train)
    
    #predictions with linear model
    y_lin_fit <- predict(fitted_linear_model, new_data = seq_train)
    # Bind the truth target variable 'ev_prop' with the predictions
    lin_result <- bind_cols(.row=1:length(seq_train$ev_prop),ev_prop = seq_train$ev_prop,lin_pred = y_lin_fit$.pred)
    #increment progress
    pb()
    ####Step 2: Train the XGBoost on the residuals of the linear fit.
    # Extract residuals
    seq_train2 <- seq_train %>%
      bind_cols(y_lin_fit %>%
                  rename(linear_pred = .pred)) %>%
      mutate(residuals = ev_prop - linear_pred) %>%
      select(-linear_pred)
    
    ### Define Linear model Workflow
    model_name2 <- names(model_specs2)[2]
    spec2 <- model_specs2[[2]]
    initial_tune2 <- initial_tunes[[2]]
    
    #define time_series cross validation
    time_series_folds2 <- time_series_cv(seq_train2,
                                    date_var = sequence_number,
                                    initial = 1,
                                    assess = 1,
                                    cumulative = TRUE,
                                    skip = 1)
    #set plan
    time_series_folds2 %>% tk_time_series_cv_plan()

    # Define the recipe for training xgboost on residuals
    residuals_recipe <- recipe(residuals ~ ., data = seq_train2) %>%
      step_rm(sequence_number, ev_prop) %>%
      update_role(lga_code, new_role = "id") %>% # Assign 'id' a role that excludes it from modeling
      step_dummy(all_nominal(), -all_outcomes()) %>%
      step_zv(all_predictors(),-all_outcomes()) 
      
    #create workflow for xgboost model
    xgb_workflow <- 
      workflow() %>%
      add_model(spec2) %>%
      add_recipe(residuals_recipe)
    
    #get parameters to tune
    bayes_param_xgb <- xgb_workflow %>% 
      extract_parameter_set_dials() %>% 
      update(mtry = finalize(mtry(), seq_train2%>%select(-c("sequence_number","ev_prop"))))
      
    # optimise model hyperparameters
    tuned_results2 <- tune_bayes(xgb_workflow,
                         resamples = time_series_folds2,
                         param_info = bayes_param_xgb,
                         initial = initial_tune2, 
                         iter = 8,
                         control = control_bayes(verbose = TRUE, verbose_iter = TRUE, no_improve = 4)) 
    # select best model        
    xgb_best_params <- select_best(tuned_results2,metric = "rmse")
    xgb_final_wf <- finalize_workflow(xgb_workflow, xgb_best_params)
    #fit model to training set  
    fitted_xgb_model <- fit(xgb_final_wf, data = seq_train2) 
      
    #predictions with linear model
    y_xgb_fit <- predict(fitted_xgb_model, new_data = seq_train2)
    # increment progress 
    pb()
    ###Step 3: evaluate training score under rolling validation
    #collect models in a list for later prediction.
    submodel_fits <- list(linear_model = fitted_linear_model,
                        xgb_model = fitted_xgb_model,
                        imp_num = i)
    # fit resamples on xgboost learning the residuals for the set linear trend prediction
    resample_results <- fit_resamples(xgb_final_wf, resamples = time_series_folds2, control = control_resamples(save_pred = TRUE))
    
    # collect resampling predictions
    fullmodel_fit_predictions <- resample_results %>%
        collect_predictions() %>%
        mutate(imp = i)
    
    # Join y_lin_fit_df with fullmodel_fit_predictions on .row
    combined_predictions <- fullmodel_fit_predictions %>%
      left_join(lin_result, by = ".row")
    
    # Sum the linear and residual predictions together to get the boosted prediction
    combined_predictions <- combined_predictions %>%
      mutate(boosted_pred = lin_pred + .pred) %>%
      select(-c("residuals","lin_pred",".pred"))
    
    #collect all rolling validation predictions for aggregation
    all_results[[i]] <- combined_predictions  
    
    #collect all models across mids
    all_models[[i]] <- submodel_fits
    pb()
  }
  all_results <- bind_rows(all_results)
  
  # aggregate predictions at each row
  agg_results <- all_results %>% group_by(id,.row) %>%
    summarise(agg_boosted_pred = mean(10^(boosted_pred) - ev_prop_shift),
              ev_prop = mean(10^(ev_prop) - ev_prop_shift),
              .groups = 'drop')
  
  return(list(agg_results,all_models))
}

with_progress({aggregate_results_tunedmodels <- tune_models(seq_train_filtered, model_specs2, initial_tunes)})
```



# Compared tuned models
rolling time-series validation results of tuned models are compared.
```{r Evaluate tuned models}
metrics_test <- metric_set(yardstick::rmse, yardstick::mae, smape)

aggregate_result <- aggregate_results_tunedmodels[[1]]
 
modelsresults <- aggregate_result %>%
  group_by(id) %>%
  metrics_test(truth = ev_prop,estimate = agg_boosted_pred) 
modelsresults <-
modelsresults %>% 
group_by(.metric) %>%
  summarise( mean_score = mean(.estimate),
             std_score = sd(.estimate),
             .groups = 'drop')
g <- modelsresults %>%
  ggplot(aes(x = "validation", y = mean_score))+
  geom_bar(stat = "identity", fill = "skyblue", alpha = 0.7)+
  geom_errorbar(aes(ymin = mean_score - std_score, ymax = mean_score + std_score), width = 0.3, color = "red3")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  facet_wrap(vars(.metric ), scales = "free") +
  labs( title = paste0("Evaluation of trained model for training window of ",window_size," and horizon of ", horizon," under rolling validation"),
        x = "Candidate models",
        y = "mean validation score")+
  theme(plot.title = element_text(size = 11))  # Change the title size
g 
```


```{r collate models}
modelset <- aggregate_results_tunedmodels[[2]]
modelsetinfo <- list(models = modelset,
                     forecast_horizon = horizon,
                     cluster_model = cluster_no)
model_collection <- append(model_collection,list(modelsetinfo))
save.image(file = "my_workspace.RData")
```

## Forming model predictions on testing data
1. Predictions are formed for each MIDS model set for cluster and forecast horizon combination based on the testing set.
2. The predictions are aggregated across the MIDS for each cluster/forecast horizon model.
3. The aggregated predictions are evaluated against the truth values.
```{r model predictions}
horizon <- 1
cluster_no <- 1
# collect testing data
with_progress({seq_data_lists <- reframe(imputed_X_train_list, imputed_X_test_list, y_train, y_test,window_size,horizon,ti_cols)})
seq_test_list <- seq_data_lists[[2]]
#Filtering based on cluster
seq_test_filtered <- filter_datasets(seq_test_list,cluster_no)

evaluate_models <- function(model_collection, test_data, cluster_no, horizon){
  results_list <- list()
  #loop through collections
  for (model_set in model_collection){
    # Identify the modelset as specified
    if(model_set$cluster_model == cluster_no && model_set$forecast_horizon == horizon){
      #Extract the modelset
      submodel_fits_list <- model_set$models
      all_predictions <- list()
      #loop through each model and mid
      for (i in 1:m_imps) {
        # extract the correct model for each MID
        sub_models <- submodel_fits_list[[i]]
        #extract the corresponding test MIDS and order by sequence number and transform variables
        seq_test <- data.frame(test_data[[i]])
        seq_test <- seq_test %>% 
          arrange(sequence_number) %>% # transform variables in dataset
          mutate(ev_prop = log(ev_prop + ev_prop_shift,10),
                 across(starts_with("population_density"),~log(. + pop_dense_shift,10)),
                 across(starts_with("neighbourhood_eV_prop"),~log(. + ev_prop_shift,10)))
        
        #### make predictions with sub-models
        # Generalised Linear model
        fitted_linear_model <- sub_models$linear_model
        trend_predictions <- predict(fitted_linear_model, new_data = seq_test)
        #xGBoost Residual model
        fitted_xgb_model <- sub_models$xgb_model
        residual_predictions <- predict(fitted_xgb_model, new_data = seq_test)
        #Boost the predictions by adding the y_preds
        boosted_predictions <- bind_cols(trend_pred = trend_predictions$.pred, res_pred = residual_predictions$.pred) %>%
          mutate(boost_pred = trend_pred + res_pred)
        # combine with id and truth value  
        predictions <- boosted_predictions %>%
          bind_cols(seq_test%>%select(c("ev_prop","lga_code","sequence_number"))) %>%
          mutate(imp = i)
        #append predictions to the list
        all_predictions[[i]] <- predictions
      }
      # combine all predictions
      combined_predictions <- bind_rows(all_predictions)
      agg_predictions <- combined_predictions %>%
        group_by(lga_code,sequence_number) %>%
        summarise(agg_pred = mean(10^(boost_pred)-ev_prop_shift),
                  ev_prop = mean(10^(ev_prop)-ev_prop_shift),
                  .groups = 'drop') %>%
        ungroup()
    }
  }
  return(list(agg_predictions,combined_predictions))
}

agg_predictions_combined <- evaluate_models(model_collection, seq_test_filtered, cluster_no, horizon)
combined_pred <- agg_predictions_combined[[2]]
model_results1 <- metrics_test(agg_predictions_combined[[1]],truth = ev_prop,estimate = agg_pred) 

g <- model_results1 %>%
ggplot(aes(x = "Evaluation Metrics", y = .estimate))+
geom_bar(stat = "identity", fill = "skyblue", alpha = 0.7)+
theme(axis.text.x = element_text(angle = 45,hjust = 1))+
facet_wrap(vars(.metric ), scales = "free") +
labs( title = paste0("Evaluation of trained model for training window of ",window_size," and horizon of ", horizon," under rolling validation"),
      x = "Candidate models",
      y = "mean validation score")+
theme(plot.title = element_text(size = 11))  # Change the title size
g
```

```{r}
submodel_fits[1]
```


